### 2017-02-17 CS798-ConvexityAndOptimization/assignments/assignment_01.tex
p6

%Q5
\section{}
Let $Q \in S^n_{++}$. We consider the quadratic norm $||x||_Q := (x^TQx)^\frac{1}{2}.$
%pa
\subsection{a}
\paragraph{}
We show that $||x||_Q$ is a norm. First since $Q\in S^n_{++}$ there exists non-singular, symmetric $V$ such that $Q = V^TV$. Observe that
$$||x||_Q = (x^TQx)^\frac{1}{2} = (x^TV^TVx)^\frac{1}{2} =  ((Vx)^T(Vx))^\frac{1}{2} = ||Vx||_2.$$
We will show that $||x||_Q$ is non-negative and equal to $0$ only when $x = 0$, that $||t x||_Q = |t|\cdot ||x||_Q$, and that triangle inequality holds.
\paragraph{Non-negativity}
Since $||\cdot||_2$ is a norm, $||x||_Q = ||Vx||_2 \geq 0$ and further $||Vx||_2 = ||x||_Q = 0$ only if $Vx = 0$, but then $x = V^{-1}0 = 0$.
\paragraph{Scaling}
Observe that $$||tx||_Q = ||V(tx)||_2 = |t|\cdot ||Vx||_2 = |t|\cdot ||x||_Q.$$
\paragraph{Triangle Inequality}
Let $x,y \in \R^n$. Then
$$||x+y||_Q = ||V(x+y)||_2 = ||Vx + Vy||_2 \leq ||Vx||_2 + ||Vy||_2 = ||x||_Q + ||y||_Q.$$
Therefore $||\cdot||_Q$ is a norm. $\blacksquare$
%pb
\subsection{b}
\paragraph{}
Let $x \in \R^n$. Then observe that
\begin{align*}
||x||_{Q^*} = \sup \{ x^Ty : ||y||_Q \leq 1\} \\
&= \sup \{x^Ty : ||Vy||_2 \leq 1\} \\
&= \sup \{x^TV^{-1}Vy : ||Vy||_2 \leq 1\} \\
&= \sup\{x^TV^{-1}z : ||z||_2 \leq 1\} &\text{assigning $z=Vy$}. 
\end{align*}
The supremum is realized when $z = (x^TV^{-1})^T = V^{-1}x$. Therefore
$$||x||_{Q^*} = x^TV^{-1}V^{-1}x = x^TQ^{-1}x = ||x||_{Q^{-1}}.$$
Thus our nice formula is $||x||_{Q^*} = ||x||_{Q^{-1}}$.$\blacksquare$
%pc
\subsection{c}
\paragraph{}
Let $\cE = \{x : x^TQx \leq 1\}.$  Rewriting we have $\cE = \{x : ||x||_Q \leq 1\} = \cB$ where $\cB$ denote the norm ball of $||\cdot||_Q$. Then we have
$$\cE^* = \{y : y^Tx \leq 1, \forall x \in \cE\} = \{y : y^Tx \leq 1, \forall x \in \cB\}  = \cB^* = \cB_*,$$
where $\cB_*$ is the norm ball of $||\cdot||_{Q^*}$, and the last equality follows from the notes. Rewriting using part $(b)$ we see that $\cB_* = \{x: x^TQ^{-1}x \leq 1\}$. That is $\cE^*$ is ellipsoid given by $Q^{-1}$.
\end{document}

>>>>

%Q5
\section{}
Let $Q \in S^n_{++}$. We consider the quadratic norm $||x||_Q := (x^TQx)^\frac{1}{2}.$
%pa
\subsection{a}
\paragraph{}
We show that $||x||_Q$ is a norm. First since $Q\in S^n_{++}$ there exists non-singular, symmetric $V$ such that $Q = V^TV$. Observe that
$$||x||_Q = (x^TQx)^\frac{1}{2} = (x^TV^TVx)^\frac{1}{2} =  ((Vx)^T(Vx))^\frac{1}{2} = ||Vx||_2.$$
We will show that $||x||_Q$ is non-negative and equal to $0$ only when $x = 0$, that $||t x||_Q = |t|\cdot ||x||_Q$, and that triangle inequality holds.
\paragraph{Non-negativity}
Since $||\cdot||_2$ is a norm, $||x||_Q = ||Vx||_2 \geq 0$ and further $||Vx||_2 = ||x||_Q = 0$ only if $Vx = 0$, but then $x = V^{-1}0 = 0$.
\paragraph{Scaling}
Observe that $$||tx||_Q = ||V(tx)||_2 = |t|\cdot ||Vx||_2 = |t|\cdot ||x||_Q.$$
\paragraph{Triangle Inequality}
Let $x,y \in \R^n$. Then
$$||x+y||_Q = ||V(x+y)||_2 = ||Vx + Vy||_2 \leq ||Vx||_2 + ||Vy||_2 = ||x||_Q + ||y||_Q.$$
Therefore $||\cdot||_Q$ is a norm. $\blacksquare$
%pb
\subsection{b}
\paragraph{}
Let $x \in \R^n$. Then observe that
\begin{align*}
||x||_{Q^*} = \sup \{ x^Ty : ||y||_Q \leq 1\} \\
&= \sup \{x^Ty : ||Vy||_2 \leq 1\} \\
&= \sup \{x^TV^{-1}Vy : ||Vy||_2 \leq 1\} \\
&= \sup\{x^TV^{-1}z : ||z||_2 \leq 1\} &\text{assigning $z=Vy$}. 
\end{align*}
The supremum is realized when $z = (x^TV^{-1})^T = V^{-1}x$. Therefore
$$||x||_{Q^*} = x^TV^{-1}V^{-1}x = x^TQ^{-1}x = ||x||_{Q^{-1}}.$$
Thus our nice formula is $||x||_{Q^*} = ||x||_{Q^{-1}}$.$\blacksquare$
%pc
\subsection{c}
\paragraph{}
Let $\cE = \{x : x^TQx \leq 1\}.$  Rewriting we have $\cE = \{x : ||x||_Q \leq 1\} = \cB$ where $\cB$ denote the norm ball of $||\cdot||_Q$. Then we have
$$\cE^* = \{y : y^Tx \leq 1, \forall x \in \cE\} = \{y : y^Tx \leq 1, \forall x \in \cB\}  = \cB^* = \cB_*,$$
where $\cB_*$ is the norm ball of $||\cdot||_{Q^*}$, and the last equality follows from the notes. Rewriting using part $(b)$ we see that $\cB_* = \{x: x^TQ^{-1}x \leq 1\}$. That is $\cE^*$ is ellipsoid given by $Q^{-1}$.
\paragraph{}
So suppose $\cE$ is a $2D$-ellipsoid with semi-axes $v_1$ and $v_2$ with corresponding side lengths $a_1$ and $a_2$. Then the spectral decomposition of $Q$ is
$$Q = \begin{bmatrix} v_1 & v_2 \end{bmatrix} \begin{bmatrix} a_1^{-1} & 0 \\ 0 & a_2^{-1} \end{bmatrix} \begin{bmatrix} v_1^T \\v_2^T\end{bmatrix}.$$
And further
$$Q^{-1} =  \begin{bmatrix} v_1 & v_2 \end{bmatrix} \begin{bmatrix} a_1 & 0 \\ 0 & a_2 \end{bmatrix} \begin{bmatrix} v_1^T \\v_2^T\end{bmatrix}$$
using that $ \begin{bmatrix} v_1 & v_2 \end{bmatrix}^T =  \begin{bmatrix} v_1 & v_2 \end{bmatrix}^{-1}$. Hence the polar ellipsoid $\cE^*$ has the same semi-axes as $\cE$, but with reciprocal side lengths.
\begin{figure}[H]
\centering
\begin{tikzpicture}
\draw (0,0) ellipse (0.5cm and 2cm);
\draw (0,0) ellipse (2cm and 0.5cm);
\end{tikzpicture}
\caption{Ellipsoid and its polar where $v_1 = e_1$ and $v_2 = e_2$ (standard basis vectors), and $a_1 = 2$ while $a_2 = \frac{1}{2}$. The polar scales $v_1$ by $\frac{1}{a_1} = \frac{1}{2}$ and scales $v_2$ by $\frac{1}{a_2} = 2$.}
\end{figure}

%Q6
\section{}
%Pa
\subsection{a}
\paragraph{}
Let $\cE_1:= \{P_1 u + q_1 : ||u||_2 \leq 1\}$ and $\cE_2:= \{P_2 u + q_2 : ||u||_2 \leq 1\}$ be given ellipsoids. We want to find a separating hyperplane $(a,b)$ for these ellipsoids, that is:
\begin{align*}
a^Tx + b &\geq 0 &\forall x \in \cE_1 \\
a^Tx+b &\leq 0 &\forall x \in \cE_2.
\end{align*}
This can be formulated as an optimization feasibility problem as
\begin{align*}
\min\ 0\\
\text{s.t.}\ a^Tx + b &\geq 0 &\forall x \in \cE_1\\
a^Tx + b &\leq 0 &\forall x \in \cE_2.
\end{align*}
This problem is equivalent to
\begin{align*}
\min\ 0\\
\text{s.t.}\ a^T(-x)&\leq b &\forall x \in \cE_1\\
a^Tx &\leq -b &\forall x \in \cE_2.
\end{align*}
Which is equivalent to
\begin{align*}
\min\ 0\\
\text{s.t.}\  
\sup\{ a^T(-x) : x \in \cE_1\}&\leq b\\
\sup\{a^Tx : x\in \cE_2\} &\leq -b.
\end{align*}
Equivalently
\begin{align*}
\min\ 0\\
\text{s.t.}\  
\sup\{ a^T(-P_1 u - q_1) : ||u||_1 \leq 1\}&\leq b\\
\sup\{a^T(P_2 u + q_2) : ||u||_2 \leq 1\} &\leq -b.
\end{align*}
Which is equivalent to
\begin{align*}
\min\ 0\\
\text{s.t.}\  
a^T(-q_1) + ||P_1a||_2 &\leq b\\
a^Tq_2 + ||P_2a||_2 &\leq -b.
\end{align*}
Setting $A_1 = P_1$, $b_1 = 0$, $c_1=q_1$, $d_1 = b$, and $A_2 = P_2$, $b_2 = 0$, $c_2 = -q_2$, and $d_2 = -b$ we obtain the exact form of a Second Order Conic Program as described in the general formulation of such problems. $\blacksquare$
\subsection{b}
\paragraph{}
We will write the Second-Order Cone Program (SOCP):
\begin{align*}
\min &f^Tx \\
\text{s.t.} ||A_ix + b_i||_2 &\leq c^T_i x + d_i, &1\leq i \leq m \\
Fx &= g,
\end{align*}
as a Semidefinite Program (SDP) in inequality form (see section $4.6.2$ of Boyd and Vandenberghe) :
\begin{align*}
\min\ &f^Tx \\
\text{s.t.}\ G + \sum_{i=1}^n x_i F_i &\preccurlyeq 0 \\
Fx &=g.
\end{align*}
Observe that it is enough to show that a constraint of the form $||Ax + b|| \leq c^Tx + d$ can be written as $G + \sum_{i=1}^n x_i F_i \preccurlyeq 0$ as then we could write each constraint $||A_ix + b_i||_2 \leq c^T_i x + d_i$ for $i\in [m]$ as $G^i + \sum_{j=1} x_j F^i_j$, and stack them as blocks as
$$G = \begin{bmatrix} G^1 & & \\ & \ddots & \\ & & G^m \end{bmatrix} \quad F_i = \begin{bmatrix} F^1_i & & \\ & \ddots & \\ & & F^m_i \end{bmatrix}$$
to obtain an equivalent single semidefinite constraint in place of $m$ semidefinite constraints.
\paragraph{Lemma 6b}
We will need the following lemma using Schur's complement: For any $x\in \R^n$ and $c\in \R$ we have $||x||_2 \leq c$ if and only if
$$X = \begin{bmatrix} c &x^T \\ x & cI\end{bmatrix} \succcurlyeq 0.$$
\paragraph{Proof of Lemma 6b}
If $c = 0$ this is trivial, so we may assume that $c\neq 0$. In this case we invoke Schur's Complement Lemma and thus
$$X \succcurlyeq 0$$
if and only if $c \geq 0$ and $tI - \frac{1}{c}xx^T \succcurlyeq 0$. Which holds if and only if $c^2I - xx^T \succcurlyeq 0$. Now the eigenvalues of $c^2I - xx^T$ are non-negative if and only if $c^2 - ||x||^2_2 \geq 0$ (since $c^2$ is the least eigenvalue of $c^2I$ and $||x^2_2$ is the largest eigenvalue of $xx^T$), which holds if and only if $||x||_2 \leq c$ as desired. $\blacksquare$
\paragraph{}
Now applying Lemma $6b$ to our SOCP cone constraint we have $||Ax +b||_2 \leq c^Tx + d$ (suppose $x\in\R^n$ and $A \in \R^{r\times n}$) if and only if
$$\begin{bmatrix} c^Tx + d & (Ax+b)^T \\ Ax + b &(c^T+d)xI \end{bmatrix} \succcurlyeq 0.$$
Let $a^1, \dots, a^r$ be the rows of $A$. Separating out we obtain:
$$\begin{bmatrix} d & b^T \\ b & dI \end{bmatrix} + \sum_{i=1}^n x_i \begin{bmatrix} c_i & a^1_i & \dots & a^r_i \\ a^1_i \\ \vdots & &c_iI\\ a^r_i \end{bmatrix} \succcurlyeq 0.$$
Thus $$G = \begin{bmatrix} d & b^T \\ b & dI \end{bmatrix}$$
and 
$$F_i = \begin{bmatrix} c_i & a^1_i & \dots & a^r_i \\ a^1_i \\ \vdots & &c_iI\\ a^r_i \end{bmatrix}$$
we have
$$G + \sum_{i=1}^n F_i \succcurlyeq 0$$
if and only if
$$||Ax+b||_2 \leq c^Tx + d.$$
Therefore we have shown that any SOCP can be written as an SDP. $\blacksquare$
\end{document}
